---
output:
  pdf_document: default
  html_document: default
---


# Practical 5 Statistical Genetics

## Antonin Rosa

## Hendrik Wellmann

### Question 1 

```{r}
library(data.table)

data <- fread("YRI6.raw", header = FALSE, sep = " ")

num_individuals <- nrow(data)
num_snps <- ncol(data) - 6

percentage_missing <- mean(is.na(data)) * 100

cat("Number of individuals:", num_individuals, "\n")
cat("Number of SNPs:", num_snps, "\n")
cat("Percentage of missing data:", percentage_missing, "%\n")
```

### Question 2

```{r}
genomic_data <- data[2:nrow(data), 7:ncol(data)]
genomic_data[, (1:ncol(genomic_data)) := lapply(.SD, as.numeric), .SDcols = 1:ncol(genomic_data)]
shared_mean <- matrix(c(NA), nrow = num_individuals, ncol = num_individuals)
shared_sd <- matrix(c(NA), nrow = num_individuals, ncol = num_individuals)
for (i in 1:num_individuals) {
  for (j in 1:num_individuals) {
    shared <- numeric(0)
    for (k in 1:nrow(genomic_data)){
    shared <- c(shared, 2 - abs(as.matrix(genomic_data[k,i, with=FALSE]) - as.matrix(genomic_data[k,j, with=FALSE])))
    }
    mean <- mean(shared)
    sd <- sd(shared)
    shared_mean[i,j] <- mean
    shared_sd[i,j] <- sd
  }
}

print(shared_mean[1:5, 1:5])
print(shared_sd[1:5, 1:5])
```

### Question 3

```{r}
p0 <- matrix(c(NA), nrow = num_individuals, ncol = num_individuals)
p2 <- matrix(c(NA), nrow = num_individuals, ncol = num_individuals)
m <- matrix(c(NA), nrow = num_individuals, ncol = num_individuals)
for (i in 1:num_individuals) {
  for (j in 1:num_individuals) {
    shared <- numeric(0)
    for (k in 1:nrow(genomic_data)){
      shared <- c(shared, 2 - abs(as.matrix(genomic_data[k,i, with=FALSE]) - as.matrix(genomic_data[k,j, with=FALSE])))
    }
    p0[i,j] <- sum(shared == 0)/length(shared)
    p2[i,j] <- sum(shared == 2)/length(shared)
    m[i,j] = 1 - p0[i,j] + p2[i,j]
  }
}
print(p0[1:5, 1:5])
print(p2[1:5, 1:5])
print(m[1:5, 1:5])
print(shared_mean[1:5, 1:5])
print(all.equal(shared_mean, m))
```

It holds because we can see that each element of m (m = 1 - p0 + p2) is equal to each element of shared_mean.

### Question 4

```{r}
plot(shared_mean, shared_sd, main = "m vs s", xlab = "m", ylab = "s", col = "blue", pch = 16)
plot(p0, p2, main = "p0 vs p2", xlab = "p0", ylab = "p2", col = "green", pch = 16)
```

#### m vs s : 

**Cluster in the Top Left:**
The clustered points in this region could indicate a group of individuals with similar genetic characteristics. This grouping might be due to close familial relationships or specific genetic similarities.

**Isolated Point in the Bottom Right:**
An isolated point suggests the presence of an individual or a small group of individuals with distinct genetic characteristics or differences from the majority. This could be due to unique genetic variations or atypical familial relationships.

#### p0 vs p2 : 

**Cluster on the Entire Bottom Part:**
The concentration of points in the bottom part of the plot suggests that the majority of individuals share similar genetic characteristics in terms of *p0* and *p2*. This could result from genetic similarities within this population.

**Isolated Point in the Top Left:**
The presence of an isolated point in the top left suggests that there is an individual or a group of individuals that stand out from the others in terms of *p0* and *p2*. This could indicate unique genetic variations or distinct features.

### Question 5

```{r}
m <- shared_mean
s <- shared_sd
family_relationship <- data[, c(3, 4)]
colors <- rep("blue", nrow(data) - 1)  
parent_offspring_indices <- which(family_relationship[, 1] > 0 | family_relationship[, 2] > 0)
colors[parent_offspring_indices - 1] <- "red"
parent <- c()
for (i in 2:nrow(data)) {
  print(data[i, 2] %in% family_relationship[[1]])
  if (data[i, 2] %in% family_relationship[[1]] | data[i, 2] %in% family_relationship[[2]]) {
    parent <- c(parent, i - 1)
  }
}
print(parent)
colors[parent] <- "red"

plot(m, s, main = "Shared Mean vs Shared SD", xlab = "Shared Mean", ylab = "Shared SD", col = colors, pch = 16)
legend("bottomleft", legend = c("Unrelated", "Parent-Offspring"), col = c("blue", "red"), pch = 16, cex = 0.8)

```

We can see that all the majority of points present in the clusters have a family relationship. It affirms the hypothesis said before in question 4.
